---
title: "ltm vs. mirt (2)"
author: "Piotr Król"
date: "18 06 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Content:

* 5x5 matrix
* Baker&Kim matrices
* 50x5 random matrix
* LSAT7
* 100X5 random matrix
* CDI items params
* CDI CAT simulations

```{r libraries, message=FALSE, warning=FALSE}
library(ltm)
library(mirt)
library(mirtCAT)
library(parallel) #Speed up simulations
library(plyr) #For laply
library(ggpubr)
```

## 5x5 matrix

Let's take again matrix (second one) from last report:

```{r }
responses <- rbind(
  c(1,0,1,0,1),
  c(0,1,0,1,0),
  c(1,0,1,0,1),
  c(0,1,0,1,0),
  c(1,1,0,0,1)
)
colnames(responses) <- paste0("item", c(1:5)) #needded for mirt
responses
```

And according to https://groups.google.com/g/mirt-package/c/yLV12DSVj2A check confidence intervals for parameters:

```{r results='hide'}
mirt_model <- mirt(responses, 1, SE=TRUE)
```

```{r }
coef(mirt_model)
```

They are in fact pretty wide. Let's look at item characteristic curves:

```{r}
plot(mirt_model, type = 'trace')
```

They don't look good. Let's look for a "numerically stable" model and check if in such a case parameters would be similar.

## Baker & Kim matrices (ch. 7)

I will take matrix from Baker & Kim (2017, p.116):

```{r}
N <- 16
J <- 10
U <- matrix(c(
0, 0, 1, 0, 0, 0, 0, 1, 0, 0,
1, 0, 1, 0, 0, 0, 0, 0, 0, 0,
1, 1, 1, 0, 1, 0, 1, 0, 0, 0,
1, 1, 1, 0, 1, 0, 0, 0, 0, 0,
0, 0, 0, 0, 1, 0, 0, 0, 0, 0,
1, 1, 0, 1, 0, 0, 0, 0, 0, 0,
1, 0, 0, 0, 0, 1, 1, 1, 0, 0,
1, 0, 0, 0, 1, 1, 0, 0, 1, 0,
1, 0, 1, 0, 0, 1, 0, 0, 1, 0,
1, 0, 0, 0, 1, 0, 0, 0, 1, 0,
1, 1, 0, 1, 1, 1, 1, 1, 1, 1,
1, 1, 1, 1, 1, 1, 1, 1, 1, 0,
1, 1, 1, 0, 1, 0, 1, 0, 0, 1,
1, 1, 1, 1, 1, 1, 1, 1, 1, 0,
1, 1, 0, 1, 1, 1, 1, 1, 1, 1,
1, 1, 1, 1, 1, 1, 1, 1, 1, 1
), nrow=N, ncol=J, byrow=TRUE)
colnames(U) <- paste0("item", c(1:J))
U
```

```{r results = 'hide'}
mirt_model <- mirt(U, 1, SE=TRUE)
```

```{r}
mirt_model
```

*mirt* fails to converge. I have probably to remove ceiling subject (16th):

```{r}
U <- U[-c(16), ]
U
```

```{r results = 'hide'}
mirt_model <- mirt(U, 1, SE=TRUE)
```

```{r}
mirt_model
```

Still fails to converge but there are no any ceiling or floor items/subjects:

```{r}
rowSums(U)
colSums(U)
```

Let's take second matrix from Baker & Kim (2017, p.117):

```{r}
N <- 16
J <- 10
U2 <- matrix(c(
0, 0, 1, 0, 0, 0, 0, 0, 0, 0,
0, 1, 1, 1, 0, 0, 0, 1, 0, 0,
1, 0, 0, 1, 0, 0, 0, 0, 0, 0,
1, 1, 1, 0, 0, 0, 1, 0, 0, 1,
1, 1, 1, 1, 0, 0, 0, 0, 0, 1,
1, 1, 1, 0, 1, 1, 0, 0, 0, 1,
1, 1, 1, 1, 1, 1, 1, 1, 1, 0,
1, 1, 1, 0, 1, 1, 1, 0, 0, 0,
0, 1, 1, 1, 1, 1, 0, 1, 1, 0,
1, 0, 1, 1, 1, 1, 1, 0, 1, 0,
1, 1, 1, 1, 1, 0, 0, 1, 0, 1,
1, 1, 1, 1, 1, 0, 1, 1, 1, 0,
1, 1, 1, 0, 1, 1, 0, 0, 0, 0,
1, 1, 1, 0, 1, 1, 1, 1, 1, 1,
1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
1, 1, 1, 1, 1, 1, 1, 1, 1, 1
), nrow=N, ncol=J, byrow=TRUE)
colnames(U2) <- paste0("item", c(1:J))
U2
```

```{r results = 'hide'}
mirt_model <- mirt(U2, 1, SE=TRUE)
```

```{r}
mirt_model
```

This also fails to converge. 

## 50x5 random matrix

Let's generate some matrix (50x5) randomly filled with 0s and 1s:

```{r eval=FALSE}
r = 50
c = 5
responses <- matrix(rbinom(r*c,1,0.5),r,c)
colnames(responses) <- paste0("item", c(1:c))
```

```{r eval=FALSE, include=FALSE}
save(responses, file = "Data/responses")
```

```{r include=FALSE}
load("Data/responses")
```

```{r}
head(responses)
```

```{r}
colSums(responses)
```

```{r}
rowSums(responses)
```

And create *mirt* model:

```{r results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE)
```

```{r}
mirt_model
```

Now succesfully converged. Let's look at parameters:

```{r}
coef(mirt_model)
```

```{r}
plot(mirt_model, type = 'trace')
```

```{r}
itemfit(mirt_model)
```

So it looks like curves are generally fitting (maybe except 1st).

Let's check for *ltm*:

```{r}
(ltm_model <- ltm(responses ~ z1, IRT.param = FALSE))
```

```{r}
plot(ltm_model)
```

So it looks like parameters are similar but slopes have opposite sign??

## LSAT7

Let's check on LSAT7 dataset:

```{r}
responses <- expand.table(LSAT7)
head(responses)
```

```{r}
dim(responses)
```

```{r results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE)
```

```{r}
coef(mirt_model)
```

```{r}
plot(mirt_model, type = 'trace')
```

Nice curves. Let's check *ltm* results:

```{r}
(ltm_model <- ltm(responses ~ z1, IRT.param = FALSE))
```

```{r}
plot(ltm_model)
```

It looks pretty similar.

## 100x5 matrix

Let's take another random matrix:

```{r eval=FALSE}
r = 100
c = 5
responses <- matrix(rbinom(r*c,1,0.5),r,c)
colnames(responses) <- paste0("item", c(1:c))
```

```{r eval=FALSE, include=FALSE}
save(responses, file = "Data/responses100")
```

```{r include=FALSE}
load("Data/responses100")
```

```{r}
head(responses)
```

```{r}
colSums(responses)
```

```{r}
rowSums(responses)
```

```{r results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE)
```

```{r}
coef(mirt_model)
```

```{r}
plot(mirt_model, type = 'trace')
```

```{r}
(ltm_model <- ltm(responses ~ z1, IRT.param = FALSE))
```

```{r}
plot(ltm_model)
```

Results from *mirt* and *ltm* are pretty similar.

## CDI data

Load already prepared data:

```{r message=FALSE}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
cdi <- read.csv("Data/cdi.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
```

Create models of CDI data:

```{r eval = FALSE}
ltm_model <- ltm(responses ~ z1, IRT.param = FALSE)
mirt_model <- mirt(responses, 1, SE = TRUE)
```

```{r include=FALSE}
load("Data/ltm_model")
load("Data/mirt_model")
```

Look at some global fit statistics both from *ltm* and *mirt*

```{r}
mirt_model
```

```{r }
ltm_summary <- summary(ltm_model)
ltm_summary$logLik
ltm_summary$AIC
ltm_summary$BIC
```

Let's look at parameters estimates from both models:

```{r,  fig.width=15}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
ltm_params <- as.data.frame(coef(ltm_model))
colnames(ltm_params) <- c("d", "z1")

mirt_params <- cbind(mirt_params, cdi)
ltm_params <- cbind(ltm_params, cdi)

mirt_plot <- ggplot(mirt_params, aes(d, a1, colour = category)) +
  geom_point() +
  xlab("Intercept") +
  ylab("Slope") +
  labs(title = "mirt") + 
  theme(legend.position = "none")

ltm_plot <- ggplot(ltm_params, aes(d, z1, colour = category)) +
  geom_point() +
  xlab("Intercept") +
  ylab("Slope") + 
  labs(title = "ltm") +
  theme(legend.position = "none")

ggarrange(mirt_plot, ltm_plot)
```

So in proper scale they look pretty similar. General order is preserved (points are coloured by categories).

Let's set the same axes for both plots:

```{r fig.width=15}
mirt_plot <- ggplot(mirt_params, aes(d, a1, colour = category)) +
  geom_point() +
  xlab("Intercept") +
  ylab("Slope") +
  labs(title = "mirt") + 
  theme(legend.position = "none") +
  xlim(-5,5) + 
  ylim(0,5)

ltm_plot <- ggplot(ltm_params, aes(d, z1, colour = category)) +
  geom_point() +
  xlab("Intercept") +
  ylab("Slope") + 
  labs(title = "ltm") +
  theme(legend.position = "none") +
  xlim(-5,5) + 
  ylim(0,5)

ggarrange(mirt_plot, ltm_plot)
```

Now it's visible that *ltm* params are somehow squished.

## Simulations

Let's check how simulations will behave with *mirt* and *ltm* params.

### 10 items

```{r}
cl <- makeCluster(detectCores()) #Speeds up simulation
mirt_obj <- generate.mirt_object(mirt_params[1:2], '2PL') 
ltm_obj <- generate.mirt_object(data.frame(a1 = ltm_params$z1, d = ltm_params$d), '2PL')

items_nr <- 10
design <- list(min_items = items_nr, max_items = items_nr)
results_mirt <- mirtCAT(mo = mirt_obj, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = design)
results_ltm <- mirtCAT(mo = ltm_obj, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = design)
```

```{r fig.width=12}
thetas <- laply(results_mirt, function(x) x$thetas)
SE_thetas <- laply(results_mirt, function(x) x$SE_thetas)
plot_mirt <- ggscatter(data.frame(thetas, SE_thetas), y = "SE_thetas", x = "thetas", 
                  xlab = "Theta", ylab = "SE", title = paste0("mirt params - average SE: ", round(mean(SE_thetas),2)))

thetas <- laply(results_ltm, function(x) x$thetas)
SE_thetas <- laply(results_ltm, function(x) x$SE_thetas)
plot_ltm<- ggscatter(data.frame(thetas, SE_thetas), y = "SE_thetas", x = "thetas",
                  xlab = "Theta", ylab = "SE", title = paste0("ltm params - average SE: ", round(mean(SE_thetas),2)))

ggarrange(plot_mirt, plot_ltm)
```

```{r fig.width=12}
raw_scores <- rowSums(responses)

thetas <- laply(results_mirt, function(x) x$thetas)
SE_thetas <- laply(results_mirt, function(x) x$SE_thetas)
plot_mirt <- ggscatter(data.frame(thetas, raw_scores), y = "raw_scores", x = "thetas",
          xlab = "Theta", ylab = "Raw score", title = "mirt")

thetas <- laply(results_ltm, function(x) x$thetas)
SE_thetas <- laply(results_ltm, function(x) x$SE_thetas)
plot_ltm <- ggscatter(data.frame(thetas, raw_scores), y = "raw_scores", x = "thetas",
          xlab = "Theta", ylab = "Raw score", title = "ltm")

ggarrange(plot_mirt, plot_ltm)
```

## 20 items

```{r}
items_nr <- 20
design <- list(min_items = items_nr, max_items = items_nr)
results_mirt <- mirtCAT(mo = mirt_obj, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = design)
results_ltm <- mirtCAT(mo = ltm_obj, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = design)
```

```{r fig.width=12}
thetas <- laply(results_mirt, function(x) x$thetas)
SE_thetas <- laply(results_mirt, function(x) x$SE_thetas)
plot_mirt <- ggscatter(data.frame(thetas, SE_thetas), y = "SE_thetas", x = "thetas", 
                  xlab = "Theta", ylab = "SE", title = paste0("mirt params - average SE: ", round(mean(SE_thetas),2)))

thetas <- laply(results_ltm, function(x) x$thetas)
SE_thetas <- laply(results_ltm, function(x) x$SE_thetas)
plot_ltm<- ggscatter(data.frame(thetas, SE_thetas), y = "SE_thetas", x = "thetas",
                  xlab = "Theta", ylab = "SE", title = paste0("ltm params - average SE: ", round(mean(SE_thetas),2)))

ggarrange(plot_mirt, plot_ltm)
```

```{r fig.width=12}
thetas <- laply(results_mirt, function(x) x$thetas)
SE_thetas <- laply(results_mirt, function(x) x$SE_thetas)
plot_mirt <- ggscatter(data.frame(thetas, raw_scores), y = "raw_scores", x = "thetas",
          xlab = "Theta", ylab = "Raw score", title = "mirt")

thetas <- laply(results_ltm, function(x) x$thetas)
SE_thetas <- laply(results_ltm, function(x) x$SE_thetas)
plot_ltm <- ggscatter(data.frame(thetas, raw_scores), y = "raw_scores", x = "thetas",
          xlab = "Theta", ylab = "Raw score", title = "ltm")

ggarrange(plot_mirt, plot_ltm)
```