---
title: "Quadratures"
author: "Piotr Król"
date: "25 06 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Content:

* ltm
  + 15 quadratures
  + 61 quadratures
* mirt
  + 15 quadratures
  + 61 quadratures
  + 151 quadratures
  + 151 quadratures with increased max number of cycles
  + 301 quadratures
  + 301 quadratures with increased max number of cycles
  + 451 quadratures
  + 601 quadratures
* summary
  + differences in mirt's logLik wrt quadratures
  + differences in mirt's 1st item param wrt quadratures

```{r libraries, message=FALSE, warning=FALSE}
library(ltm)
library(mirt)
library(ggpubr)
library(beepr)
```

```{r message=FALSE}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
```

## *ltm*

### *ltm* with default 15 quadrature points

```{r eval=FALSE}
ltm_model <- ltm(responses ~ z1, IRT.param = FALSE)
```

```{r include=FALSE}
load("Data/ltm_model")
```

```{r}
summary(ltm_model)$logLik
```

```{r}
ltm_params <- as.data.frame(coef(ltm_model))
head(ltm_params)
```

### *ltm* with 61 quadrature points

```{r eval=FALSE}
ltm_model <- ltm(responses ~ z1, IRT.param = FALSE, control = list(GHk = 61))
```

```{r eval=FALSE, include=FALSE}
save(ltm_model, file = "Data/ltm_model_61q")
```

```{r include=FALSE}
load("Data/ltm_model_61q")
```

```{r}
summary(ltm_model)$logLik
```

```{r}
ltm_params <- as.data.frame(coef(ltm_model))
head(ltm_params)
```

## *mirt*

### *mirt* with 15 quadrature points

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE, quadpts = 15)
```

```{r eval=FALSE, include=FALSE}
save(mirt_model, file = "Data/mirt_model_15q")
```

```{r include=FALSE}
load("Data/mirt_model_15q")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

### *mirt* with default 61 quadrature points

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE)
```

```{r include=FALSE}
load("Data/mirt_model")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

### *mirt* with 151 quadrature points

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE, quadpts = 151)
beep(sound=8)
```

```{r eval=FALSE, include=FALSE}
save(mirt_model, file = "Data/mirt_model_151q")
```

```{r include=FALSE}
load("Data/mirt_model_151q")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

### *mirt* with 151 quadrature points and 3000 max number of cycles

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE, quadpts = 151, technical=list(NCYCLES=3000))
beep(sound=8)
```

```{r eval=FALSE, include=FALSE}
save(mirt_model, file = "Data/mirt_model_151q_3000c")
```

```{r include=FALSE}
load("Data/mirt_model_151q_3000c")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

### *mirt* with 301 quadrature points

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE, quadpts = 301)
```

```{r eval=FALSE, include=FALSE}
save(mirt_model, file = "Data/mirt_model_301q")
```

```{r include=FALSE}
load("Data/mirt_model_301q")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

### *mirt* with 301 quadrature points and 3000 max number of cycles

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE, quadpts = 301, technical=list(NCYCLES=3000))
```

```{r eval=FALSE, include=FALSE}
save(mirt_model, file = "Data/mirt_model_301q_3000c")
```

```{r include=FALSE}
load("Data/mirt_model_301q_3000c")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

So clearly it's worth to increase maximum number of cycles.

### *mirt* with 451 quadrature points

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE, quadpts = 451, technical=list(NCYCLES=3000))
```

```{r eval=FALSE, include=FALSE}
save(mirt_model, file = "Data/mirt_model_451q")
```

```{r include=FALSE}
load("Data/mirt_model_451q")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

### *mirt* with 601 quadrature points

```{r eval=FALSE, results = 'hide'}
mirt_model <- mirt(responses, 1, SE=TRUE, quadpts = 601, technical=list(NCYCLES=10000))
```

```{r eval=FALSE, include=FALSE}
save(mirt_model, file = "Data/mirt_model_601q")
```

```{r include=FALSE}
load("Data/mirt_model_601q")
```

```{r}
mirt_model
```

```{r}
mirt_params <- as.data.frame(coef(mirt_model, simplify = TRUE)$items)[1:2]
head(mirt_params)
```

## Summary

Let's check how logLik and parameters for 1st item were changing with number of quadratures for *mirt* package:

```{r}
df <- data.frame(
  quadratures = c(15, 61, 151, 301, 451, 601),
  logLik = c(-472591.2, -463630.5, -462896.9, -462860.8, -462860.681, -462860.656),
  a1_1 = c(0.4435, 0.9245, 1.1699, 1.2304, 1.2310, 1.2312),
  d_1 = c(0.4609, 0.3187, 0.2477, 0.4003, 0.4041, 0.4059)
)
df
```

Some stabilization can be seen from ~301 quadratures

```{r}
ggplot(df, aes(quadratures, logLik)) +
  geom_point() +
  geom_line()
```

```{r}
ggplot(df, aes(quadratures)) +
  geom_line(aes(y = a1_1, colour = "Slope")) + 
  geom_line(aes(y = d_1, colour = "Intercept")) +
  xlab("Quadratures") +
  ylab("Value of parameter") +
  labs(title = "Parameters for Item 1")
```

