---
title: "Simulations with different number of quadratures"
author: "Piotr Król"
date: "30 07 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Content

Simulations with Kachergis stop criteria on:

* Model with default 61 quadratures
* Model with 601 quadratures
* Model with 451 quadratures

```{r libraries, warning=FALSE, message=FALSE}
library(mirt)
library(mirtCAT)
library(parallel)
library(plyr)
```

```{r data loading, message=FALSE}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
load("Data/9_simulations&quadratures.RData")
```

# Model with default 61 quadratures

```{r}
mod61q
```
```{r eval=F}
params <- as.data.frame(coef(mod61q, simplify = TRUE)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
sim_results_61q <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.15, min_items = 25, max_items = 50))
fscores61q <- fscores(mod61q, method = "MAP")
```

```{r}
sim_results <- sim_results_61q
thetas <- laply(sim_results, function(x) x$thetas)
fscores <- fscores61q
```

#### Correlation

```{r}
round(cor(thetas, as.numeric(fscores)), 3)
```

#### Average SE

```{r}
SEthetas <- laply(sim_results, function(x) x$SE_thetas)
round(mean(SEthetas),3)
```

#### Mean number of items used in each test

```{r}
items_answered <- laply(sim_results, function(x) length(x$items_answered))
round(mean(items_answered), 1)
```
# Model with 601 quadratures

```{r}
mod601q
```
```{r eval=F}
params <- as.data.frame(coef(mod601q, simplify = TRUE)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
sim_results_601q <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.15, min_items = 25, max_items = 50))
fscores601q <- fscores(mod601q, method = "MAP")
```

```{r}
sim_results <- sim_results_601q
thetas <- laply(sim_results, function(x) x$thetas)
fscores <- fscores601q
```

#### Correlation

```{r}
round(cor(thetas, as.numeric(fscores)), 3)
```

#### Average SE

```{r}
SEthetas <- laply(sim_results, function(x) x$SE_thetas)
round(mean(SEthetas),3)
```

#### Mean number of items used in each test

```{r}
items_answered <- laply(sim_results, function(x) length(x$items_answered))
round(mean(items_answered), 1)
```

# Model with 451 quadratures

```{r}
mod451q
```

```{r eval=F}
params <- as.data.frame(coef(mod451q, simplify = TRUE)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
sim_results_451q <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.15, min_items = 25, max_items = 50))
fscores451q <- fscores(mod451q, method = "MAP")
```

```{r}
sim_results <- sim_results_451q
thetas <- laply(sim_results, function(x) x$thetas)
fscores <- fscores451q
```

#### Correlation

```{r}
round(cor(thetas, as.numeric(fscores)), 3)
```

#### Average SE

```{r}
SEthetas <- laply(sim_results, function(x) x$SE_thetas)
round(mean(SEthetas),3)
```

#### Mean number of items used in each test

```{r}
items_answered <- laply(sim_results, function(x) length(x$items_answered))
round(mean(items_answered), 1)
```