---
title: "Early-stopping CAT simulations"
author: "Piotr Król"
date: "09 08 2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

```{r libraries, warning=F, message=F}
library(ltm)
library(mirt)
library(mirtCAT)
library(parallel)
library(plyr)
library(ggpubr)
```

```{r data loading, eval=F}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
load("Data/mod_r")
load("Data/items_to_remove")
responses <- responses[, -items_to_remove]
```

```{r data loading 2, include=F}
load("Data/early_stop_CAT.RData")
```

```{r model}
mod <- mod_r
rm(mod_r)
mod
```
Calculate Cronbach's alpha for responses:

```{r}
cronbach.alpha(responses)
```

Prepare full scores, params and mirt object for simulations:

```{r preparation, eval=F}
fscores <- as.data.frame(fscores(mod, method = "MAP", full.scores.SE = T)) #Obtain full scores
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2] #Prepare params
mo <- generate.mirt_object(params, '2PL') #Prepare mirt object
```
```{r}
ggplot(fscores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Standard error (SE)") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey")
)
```

```{r}
summary(fscores$SE_F1)
```

Make early-stopping simulation with desired SE = 0.1:

```{r simulation, eval=F}
#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
results_SE1 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1))
save.image("Data/early_stop_CAT.RData")
```


```{r}
sim_results <- results_SE1

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used

paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused:", unused)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 50, 75, 100, 666, 667), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 50", "51 - 75", "76 - 100", "101 - 666", "667"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of subjects (%)") +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

Make early-stopping simulation with desired SE = 0.15:

```{r simulation 2, eval=F}
#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
results_SE15 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.15))
save.image("Data/early_stop_CAT.RData")
```


```{r}
sim_results <- results_SE15

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used

paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused:", unused)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 50, 75, 100, 666, 667), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 50", "51 - 75", "76 - 100", "101 - 666", "667"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of subjects (%)") +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

```{r fig.width=12, fig.height=8}
#Prepare cuts
cuts <- cut(tests_lengths, breaks = c(0, 50,  667), labels = c("1 - 50", "51 - 667"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of tests (%)") +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

Make early-stopping simulation with desired SE = 0.15:

```{r simulation 2, eval=F}
#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
results_R <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, min_items = 10, max_items = 50))
save.image("Data/early_stop_CAT.RData")
save(results_R, file="Data/results_R")
```


```{r}
sim_results <- results_R

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
SEs <- laply(sim_results, function(x) x$SE_thetas)
meanSE <- round(mean(SEs), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used

paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused:", unused)
```
Check correlations by age groups:

* [18,21)

```{r}
round(cor(thetas[which(responses_demo$months < 21)], fscores$F1[which(responses_demo$months < 21)]), 3)
round(1 - mean(SEs[which(responses_demo$months < 21)])**2, 3)
```

* [21,24)

```{r}
round(cor(thetas[which(responses_demo$months < 24 & responses_demo$months >= 21)], fscores$F1[which(responses_demo$months < 24 & responses_demo$months >= 21)]), 3)
round(1 - mean(SEs[which(responses_demo$months < 24 & responses_demo$months >= 21)])**2, 3)
```

* [24,27)

```{r}
cor(thetas[which(responses_demo$months < 27 & responses_demo$months >= 24)], as.numeric(fscores)[which(responses_demo$months < 27 & responses_demo$months >= 24)])
```

* [27,30)

```{r}
cor(thetas[which(responses_demo$months < 30 & responses_demo$months >= 27)], as.numeric(fscores)[which(responses_demo$months < 30 & responses_demo$months >= 27)])
```

* [30,33)

```{r}
cor(thetas[which(responses_demo$months < 33 & responses_demo$months >= 30)], as.numeric(fscores)[which(responses_demo$months < 33 & responses_demo$months >= 30)])
```

* [33,36]

```{r}
round(cor(thetas[which(responses_demo$months <= 36 & responses_demo$months >= 34)], fscores$F1[which(responses_demo$months <= 36 & responses_demo$months >= 34)]), 3)
round(1 - mean(SEs[which(responses_demo$months <= 36 & responses_demo$months >= 34)])**2, 3)
```

Plot theta vs. test length:

```{r}
ggplot(data.frame(thetas = thetas, lengths = tests_lengths), aes(thetas, lengths)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Number of items administered") +
  ggtitle("Tests lengths for SE = 0.1 stop criterion") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey")
)
```

Make early-stopping simulation with different maximum number of items administered:

```{r simulations, eval=F}
lengths <- c(5, 10, 15, 25, 35, 50, 75, 100)
results_SEXX <- data.frame(mean_length = NA, median_length = NA, cor = NA, meanSE = NA, rel = NA, unused = NA)
cl <- makeCluster(detectCores()) #Speeds up simulations - computation in parallel

for (n in lengths){
  
  #Make simulation
  sim_results <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = XX, max_items = n))
  
  #Obtain mean test length
  tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
  mean_length <- round(mean(tests_lengths), 1)
  
  #Obtain median test length
  median_length <- round(median(tests_lengths), 1)
  
  #Obtain thetas
  thetas <- laply(sim_results, function(x) x$thetas)
  
  #Get correlation of thetas with full scores
  cor <- round(cor(thetas, fscores$F1), 3)
  
  #Get mean SE
  meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)
  
  #Get reliability
  rel <- round(1 - meanSE**2, 3)
  
  #Get number of unused items
  raw_responses <- laply(sim_results, function(x) x$raw_responses)
  items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
  unused <- nrow(params) - items_used
  
  #Update data frame with results
  results_SEXX <- rbind(results_SEXX, c(mean_length, median_length, cor, meanSE, rel, unused))
  
}

results_SEXX <- na.omit(results_SEXX)
row.names(results_SEXX) <- lengths
save.image("Data/early_stop_CAT.RData")
```

```{r}
#results_fixed_length
```