---
title: "Hebrew CDI:WG Comprehension analysis based on P. Król's code"
author: "Grzegorz Krajewski"
date: "19 07 2022"
output: pdf_document
---

**Based on P. Król's excellent work on Polish data :)**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

As a preparation we first load required libraries and functions:

```{r libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(beepr)
library(plyr) # We need it for laply()
library(tidyverse) # We use it to process the data from the csv
library(mirt)
library(mirtCAT)
library(parallel)
library(ggpubr)
source(paste0(getwd(),"/Functions_new/simulations.R"))
# source(paste0(getwd(),"/Functions/get_first_item_responses.R")) Not sure what we need it for
source(paste0(getwd(),"/Functions_new/misfits_removal.R"))
source(paste0(getwd(),"/Functions_new/prepare_input_files.R"))
source(paste0(getwd(),"/Functions_new/get_start_thetas.R"))
```

# Data preparation

Then we load data from the local file,
transform to right format etc.

We split the responses from the provided scores.

```{r}
read_csv("Data/hebrew-wg.csv", skip = 2, na = character()) -> responses_demo
if_else(responses_demo$sex == "Female", "girl",
        if_else(responses_demo$sex == "Male", "boy", responses_demo$sex)) -> responses_demo$sex

scores <- select(responses_demo, ! starts_with("p1_"))
colnames(scores)[3:4] <- c("comprehension_from_file", "production_from_file")

# Comprehension:
comprehension <- select(responses_demo, 1 : last_col(2)) %>%
     mutate(across(starts_with("p1_"), ~ if_else(.x != "", 1, 0)))
rowSums(select(comprehension, starts_with("p1_"))) -> scores$comprehension_sum

# Production:
production <- select(responses_demo, 1 : last_col(2)) %>%
     mutate(across(starts_with("p1_"), ~ if_else(.x == "produces", 1, 0)))
rowSums(select(production, starts_with("p1_"))) -> scores$production_sum

# Comparing scores from file with summed responses:
write_csv(scores, "Data/scores_hebrew_wg.csv")

# Around here we might want to change column names to make them consistent with Piotr's code
# (age? sex?). Or change the names in subsequent code to make it consistently English.

```

**Warning: Scores provided in the file are different from summed**
**responses from the same file (both comp and prod)!**

Scores from the file are always greater than sums (or the same),
which suggests they were calculated on a larger set of items.
Max for comprehension from the file is `r max(scores$comprehension_from_file)`.
Max for production from the file is `r max(scores$production_from_file)`.

```{r}
mean(scores$comprehension_from_file - scores$comprehension_sum >= 0)
mean(scores$production_from_file - scores$production_sum >= 0)
```

Here we pick the comprehension data...

```{r}
comprehension -> responses_demo
# production -> responses_demo
```

...and here we read the list of items (English, Hebrew, and IDs):

```{r}
read_csv("Data/hebrew-wg.csv", n_max = 3, col_names = FALSE, col_select = 3:last_col(2)) %>%
  t %>% as.data.frame() -> cdi
colnames(cdi) <- c("English", "Hebrew", "position")

# TODO (later, since they are not unique now; also, maybe not here):
# Here we can switch the names of items for the analyses between IDs, English, and hopefully Hebrew:
# colnames(responses_demo)
```

Some English labels are duplicated as are
some Hebrew labels too.
So for the time being we work on IDs, which are fully unique.

Here we prepare the responses matrix (columns starting with "p1_").

```{r}
responses <- as.matrix(select(responses_demo, starts_with("p1_")))
print(paste0("There are ", nrow(responses), " parents who answered to ", ncol(responses),
             " items and there are ", length(which(is.na(responses))), " missing responses."))
```

# Model creation

For the model creation
we first decide on the optimal number of quadrature points.
We do it iteratively by fitting models with increasing numbers and
comparing their fit. We start with the default and increase it in
a number of arbitrary, but hopefully not unreasonable, steps.

We also compare the parameter estimates for the first item,
just out of curosity: as the goodness of fit stabilises, so
should the estimates.

*It may take some time! Switch eval on only if you really need to recalculate.*

```{r eval=FALSE}
quadpts <- c(61, 151, 301, 451, 601, 751)
quadpts_comp <- sapply(quadpts, function(x) {
     cat(paste0("\n -- ", "quadpts: ", x, " -- \n"))
     mirt(responses, 1, SE=TRUE, quadpts = x, technical = list(NCYCLES = 100000)) -> m
     list(quadpts = x, conv = m@OptimInfo$converged, logLik = m@Fit$logLik,
          item1pars = coef(m, simplify = T)$items[1, 1:2])
})
```

Here are the fits (*if eval is on*):

```{r eval=FALSE}
quadpts_comp
```

And here are the parameter estimates for the first item
(*if eval is on*):

```{r eval=FALSE}
sapply(1:ncol(quadpts_comp), function(x) quadpts_comp[,x]$item1pars)
```

Based on the above we set the number of quadrature points to 301.

```{r}
quadtp <- 301
```

We then iteratively fit the model and remove any misfits until we have fits for all items.

*It may take some time!*

We save the model to a file to save time in the future
(if you want to refit the model, switch the eval on):

```{r eval=FALSE}
output <- misfits_removal(responses=responses, quadtps=quadtp, NCYCLES=100000, p=0.001,
                          output_file = "Data/hebrew_wg.RData")
```

We load the model from a file (comment out if just fitted above)
and print a basic summary:

```{r}
load("Data/hebrew_wg.RData")
mod <- output[[1]]
items_removed <- output[[2]]
rm(output) # to save space

mod
```

## Items fit

No items were misfits in the process of model fitting.

```{r}
items_removed
```

No need to remove anything from the response matrix or the list of items.

```{r}
items_to_remove_ids <- as.numeric(laply(unlist(items_removed), function(x) substr(x, 5, nchar(x))))
cdi[which(cdi$number.ws %in% items_to_remove_ids), ]
```

```{r}
if(length(items_to_remove_ids)) {
  responses_r <- responses[, -which((colnames(responses) %in% unlist(items_removed)))]
  cdi_r <- cdi[-which(cdi$number.ws %in% items_to_remove_ids), ]
  if (nrow(cdi_r) == ncol(responses_r)) print(paste0(ncol(responses_r), " items left after removal."))
} else{
  responses_r <- responses
  cdi_r <- cdi
}
```

## Local dependence

*Check and explore the following code in the future (also compare with Kachergis et al., 2022)*

```{r eval=FALSE}
# I can't make the following code hide the results...
residuals <- residuals(mod, df.p=FALSE) # TODO: explore other parameters?
save(residuals, file = "Data/hebrew_wg_comp_residuals.RData")
```

```{r}
load("Data/hebrew_wg_comp_residuals.RData")
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA # upper diagonal contains signed Cramers V coefficients
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

```{r}
summary(as.vector(residuals_up))
```

The most highly mutually dependent positions:

```{r}
indexes <- arrayInd(which(residuals_up > 0.3), dim(residuals))
data.frame(pos1 = cdi_r[indexes[, 1], "position"], pos2 = cdi_r[indexes[, 2], "position"], CramersV = residuals_up[indexes])
```

But no pair reaches the threshold of 0.5 so we don't remove anything.

# Model exploration

## IRT items parameteres

```{r}
params_IRT <- as.data.frame(coef(mod, simplify = TRUE, IRTpars = TRUE)$items)[1:2]
cdi_r <- cbind(params_IRT, cdi_r)
```

Scatterplot of difficulty by discrimination
(*we use English labels for my sake, but could easily change to Hebrew as well;
also, if we had categories we could add them as colours*):

```{r fig.width=12, fig.height=8}
# ggplot(params_cdi, aes(b, a, colour=category, label=position)) +
ggplot(cdi_r, aes(b, a, label=English)) +
  geom_text(check_overlap = T) +
  geom_point(alpha = 0.3) +
  xlab("Difficulty") +
  ylab("Discrimination") +
#  labs(colour = "Category") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

## Ability vs. raw scores

We estimate ability levels for children from the sample based on the model.

```{r}
scores <- cbind(scores, fscores(mod, method = "MAP", full.scores.SE = TRUE))
```

*Explore the failed convergence? Does it have something to do with
the chosen MAP method?*

Plot the estimated ability level against raw scores:

```{r}
ggplot(scores, aes(comprehension_sum, F1)) +
  geom_point() +
  ylab("Ability level (theta)") +
  xlab("Raw score (sum of checked items)") +
  labs(title = "Hebrew CDI:WG Comprehension") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

The scatterplot above shows that IRT estimates get "normalised",
compared to raw scores, as can be seen on the following histograms
(flat distribution for raw scores vs. bell-like distribution for *theta* estimates).

```{r}
# Should be in two separate chunks perhaps?
hist(scores$comprehension_sum, xlab = "Raw sums", main = "")
hist(scores$F1, xlab = "Estimated ability levels", main = "")
```


## Ability vs. SE

Plot the ability level against SE:

```{r fig.width=12, fig.height=8}
ggplot(scores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (theta)") +
  ylab("Standard error (SE)") +
  labs(title = "Hebrew CDI:WG Comprehension") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

The scatterplot above shows increasing errors at both extremes of the ability level.
Perhaps due to floor and ceiling effects. Errors higher than 0.1 highly unfrequent though:

```{r}
hist(scores$SE_F1, main = "Histogram of SE", xlab = "SE")
```

## SE and ability level vs. age

The graph below shows that higher error rates are at and close to age range limits
(which is sort of obvious, given the relation between theta and age).
But it also shows that high outliers are at every age.

```{r fig.width=12, fig.height=8 }
ggplot(data.frame(age = as.factor(responses_demo$age), SE = scores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab("Standard error of ability estimations") +
  labs(title = "CDI: WS") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
  )
```

So let's see the relation between ability level and age (and first between raw scores and age).
We add SE as colour.

```{r}
ggplot(scores) +
  geom_point(aes(age, comprehension_sum))
ggplot(scores) +
  geom_point(aes(age, F1, col = SE_F1)) +
  scale_fill_gradient(low="green", high="red", aesthetics = "col") +
  labs(y = "ability level", col = "SE")
```


# Simulations

```{r}
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
```

## Whole pool for everyone

```{r eval=F}
message(paste("Starting simulation @ ", Sys.time()))
results_all <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI",
                       local_pattern = responses_r, cl = cl, design = list(min_items = ncol(responses_r)))
# results_all1 <- mirtCAT(mo = mod, method = "MAP", criteria = "MI", start_item = "MI",
#                       local_pattern = responses_r, cl = cl, design = list(min_items = ncol(responses_r)))
save(results_all, file = "Data/hebrew_wg_comp_sim_all.Rdata")
message(paste("Finished @ ", Sys.time()))
beep()
```

```{r}
load("Data/hebrew_wg_comp_sim_all.Rdata")
report_sim_results(results_all, select(scores, contains("F1")))
```

Based on the whole pool simulation we can examine decreasing average SE with each item
(we plot it up to 100 items: no need for more).

```{r fig.width=12, fig.height=8, warning=FALSE}
SE_history <- laply(results_all, function(x) x$thetas_SE_history)
meanSE_item <- as.data.frame(colMeans(SE_history))
meanSE_item$item <- row.names(meanSE_item)
colnames(meanSE_item) <- c("mean_SE", "item")
meanSE_item$item <- as.numeric(meanSE_item$item)

ggline(meanSE_item, x = "item", y = "mean_SE", numeric.x.axis = TRUE) +
  xlab ("Test length") +
  ylab ("Mean SE") +
  labs(title = "Hebrew CDI:WG Comp") +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  ) + 
  scale_y_continuous(breaks = c(0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)) + 
  scale_x_continuous(breaks = seq(0, nrow(params), by=10), limits = c(1, 100))
# Warning means we skip 324 items (because we plot the first 100 only).
```

## SE = 0.1 stop criterion

The plot above shows that we need quite a number of items (ca. 70) to reach SE=.1 on average.
Let's try a simulation.

```{r eval=F}
results_SE1 <- sim_se(0.1, file = "Data/hebrew_wg_comp_sim_se1.RData")
```

```{r}
load("Data/hebrew_wg_comp_sim_se1.Rdata")
results_SE1 <- results
report_sim_results(results_SE1, select(scores, contains("F1")))
```

```{r fig.width=12, fig.height=8}
plot_length(results_SE1, "Hebrew CDI:WG Comp", 0.1)
```

## SE = 0.15 stop criterion

```{r eval=F}
results_SE15 <- sim_se(0.15, file = "Data/hebrew_wg_comp_sim_se15.RData")
```

```{r}
load("Data/hebrew_wg_comp_sim_se15.Rdata")
results_SE15 <- results
report_sim_results(results_SE15, select(scores, contains("F1")))
```

```{r fig.width=12, fig.height=8}
plot_length(results_SE15, "Hebrew CDI:WG Comp", 0.15)
```

## Different start thetas

```{r eval=F, message=FALSE, echo=FALSE}
start_thetas <- get_start_thetas(scores, "age", "sex")
fscores_aggr <- start_thetas[[1]]
start_thetas <- start_thetas[[2]]
```

TODO: Simulations with different start thetas when stop criterion would be specified

# Input files preparation

```{r eval=F}
prepare_input_files(params, cdi_r$Hebrew, fscores_aggr,
                    question = "Does your child understand:", group = "comp", id = cdi_r$position,
                    file1 = "items-HebEng-WG-comp.csv", file2 = "startThetas-Hebrew-WG-comp.csv")
```



