---
title: "Items preselection"
author: "Piotr Król"
date: "11 08 2021"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

```{r libraries, warning=F, message=F}
library(mirt)
library(ggpubr)
library(plyr)
library(CTT)
```

```{r data loading, eval=F}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
cdi <- read.csv("Data/cdi.csv", encoding = "UTF-8")
```

```{r data loading 2, include=F}
load("Data/items_preselection.RData")
```

Calculate correlation of responses with age, CTT parameters and distance from (0,0):

```{r ctt calc, eval=F}
cor <- apply(responses, 2, FUN = function(x) cor(x, responses_demo$months))
ctt_analysis <- itemAnalysis(responses)
ctt_params <- data.frame(dffclt = ctt_analysis$itemReport$itemMean, dscrmn = ctt_analysis$itemReport$pBis, cor = cor)
ctt_params$dist <- sqrt(ctt_params$dscrmn^2 + ctt_params$cor^2) 
ctt_params <- cbind(ctt_params, cdi)
save.image("Data/items_preselection.RData")
```

```{r ctt}
head(ctt_params)
```

Select best 250 items and visualize:

```{r selection}
ctt_params_s <- ctt_params
ctt_params_s[ctt_params_s$number.ws %in% ctt_params[order(ctt_params$dist, decreasing = T), ][1:250, "number.ws"], "selected"] <- "Yes"
ctt_params_s[is.na(ctt_params_s$selected), "selected"] <- "No"

ggplot(ctt_params_s, aes(cor, dscrmn, colour=selected, label=position)) +
  geom_point(alpha = 0.3) +
  xlab("Correlation of age with item responses") +
  ylab("Discrimination") +
  labs(colour = "Selected") +
  scale_color_manual(values = c("Black", "Green")) + 
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```
Create model using only best items:

```{r eval=F}
mod_250 <- mirt(data = responses[, row.names(ctt_params_s[ctt_params_s$selected == "Yes", ])], model = 1, SE = TRUE)
```

```{r}
mod_250
```
Try increase quadratures:

```{r eval=F}
mod_250_iq <- mirt(data = responses[, row.names(ctt_params_s[ctt_params_s$selected == "Yes", ])], model = 1, SE = TRUE, quadpts = 101)
```
```{r}
mod_250_iq
```
Increase quadratures once more:

```{r eval=F}
mod_250_iq2 <- mirt(data = responses[, row.names(ctt_params_s[ctt_params_s$selected == "Yes", ])], model = 1, SE = TRUE, quadpts = 151, technical = list(NCYCLES = 10000))
```
```{r}
mod_250_iq2
```

...and once more:

```{r eval=F}
mod_250_iq3 <- mirt(data = responses[, row.names(ctt_params_s[ctt_params_s$selected == "Yes", ])], model = 1, SE = TRUE, quadpts = 301, technical = list(NCYCLES = 10000))
```
Sice difference in Log-Lik is minimal I select mod_250_iq3 for further analysis:

```{r}
mod <- mod_250_iq3
mod
```
Check items fit:

```{r eval=F}
itemfit <- itemfit(mod, method = "MAP")
```

```{r}
length(which(itemfit$p.S_X2 < 0.001))
```
Check LD:

```{r eval=F}
residuals <- residuals(mod)
residuals[lower.tri(residuals)] <- NA
```

```{r}
hist(residuals, main="Histogram of signed Cramer's V coefficients", xlab= "Cramers V")
```

```{r}
summary(as.vector(residuals))
```

Make simulation with Kachergis et al. (2021) recommended settings:

```{r simulation, eval=F}
#Obtain full scores
fscores <- as.data.frame(fscores(mod, method = "MAP"))

#Prepare params
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2]

#Prepare mirt object
mo <- generate.mirt_object(params, '2PL') 

#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
sim_results <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses[, row.names(ctt_params_s[ctt_params_s$selected == "Yes", ])], cl = cl, design = list(min_SEM = 0.15, min_items = 25, max_items = 50))

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used_nr <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used_nr
```

```{r}
paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused items:", unused)
```
```{r eval=F}
load("Data/mod_r")
fscores_full <- as.data.frame(fscores(mod_r, method = "MAP"))
```
Check correlation with fscores from full model:

```{r}
cor(thetas, fscores_full)
```

Check correlation of fscores from full and restricted model:

```{r}
cor(fscores, fscores_full)
```

Check how many of items used in preliminary CAT simulation are in currently selected 250 items:

```{r}
load("Data/preliminaryCAT_items_used")
table(preliminaryCAT_items_used %in% ctt_params_s[ctt_params_s$selected == "Yes", "number.ws"] )
```

