---
title: "Item usage in CAT"
author: "Piotr Król"
date: "01 09 2021"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Content:

* Items parameters and category belonging
* Frequency of items and categories in CAT simulation

```{r libraries, warning=F, message=F}
library(mirtCAT)
library(parallel)
library(ggpubr)
library(plyr)
```

```{r data preparation}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
cdi <- read.csv("Data/cdi.csv", encoding = "UTF-8")
load("Data/items_to_remove")
load("Data/params")
load("Data/fscores")
load("Data/mod_r")
load("Data/start_thetas")
responses <- responses[, -items_to_remove]
cdi <- cdi[!cdi$number.ws %in% items_to_remove, ]
source(paste0(getwd(),"/get_sim_results.R"))
```

```{r ready report data, include=F}
load("Data/items_usage.RData")
```

```{r }
mod_r
```

# Items parameters and category belonging

```{r eval=F}
params_IRT <- as.data.frame(coef(mod_r, simplify = T, IRTpars = T)$items)[1:2]
```

```{r fig.width=12, fig.height=8}
params_cdi <- cbind(params_IRT, cdi)
ggplot(params_cdi, aes(b, a, colour=category, label=position)) +
  geom_text(check_overlap = T) +
  geom_point(alpha = 0.3) +
  xlab("Difficulty") +
  ylab("Discrimination") +
  labs(colour = "Category") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

# Frequency of items and categories in CAT simulation

Make simulation with recommended settings and starting thetas decreased by 2:

```{r eval=F}
sim_results <- mirtCAT(
    mo = generate.mirt_object(params, '2PL'),
    method = "MAP",
    criteria = "MI",
    start_item = "MI",
    local_pattern = responses,
    cl = makeCluster(detectCores()),
    design = list(min_items = 10, min_SEM = 0.1, max_items = 50, thetas.start = start_thetas - 2)
  )
```

Check frequency of occurence of particular items:

```{r fig.width=12, fig.height=8}
#Prepare cuts
raw_responses <- laply(sim_results, function(x) x$raw_responses)
freq <- apply(raw_responses, 2, function(x) length(which(!is.na(x))))/nrow(responses)*100
cuts <- cut(freq, breaks = c(0, 0.001, 10, 20, 30, 51, 100), labels = c("0 (unused)", "0 - 10", "10 - 20", "20 - 30", "30 - 51", "93.3"), include.lowest = TRUE)

#Plot
ggplot(data.frame(table(cuts)), aes(x = cuts, y = Freq)) +
  xlab("Frequency of use (%)") +
  ylab("Number of items") +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

```{r}
params_cdi$freq <- freq
params_cdi[order(params_cdi$freq, decreasing = T), ][1:10, ]
```

```{r eval=F, include=F}
#Table for thesis
table <- params_cdi[order(params_cdi$freq, decreasing = T), ][1:10, ]
row.names(table) <- NULL
table$freq <- round(table$freq, 1)
table$dscrmn <- round(table$a, 2)
table$dffclt <- round(table$b, 2)
table$a <- NULL
table$b <- NULL
table$number.wg <- NULL
table$number.ws <- NULL
write.table(table, file="table.txt", quote = F, row.names = F)
```

Items colored by frequency:

```{r fig.width=12, fig.height=8}
ggplot(params_cdi[!params_cdi$position %in% c('szukać'), ], aes(b, a, label=position, colour=freq)) + #Remove 'szukać' since it has 93.3% frequency
  geom_text(aes(label=ifelse(freq > 0, as.character(position), " ")), check_overlap = T) +
  geom_text(aes(label=ifelse(freq > 30, as.character(position), " ")), check_overlap = F) +
  geom_point(alpha = 0.3) +
  annotate("text", x = params_cdi[params_cdi$number.ws == 508, "b"] - 0.05, y = params_cdi[params_cdi$number.ws == 508, "a"], colour = "black", label = "szukać") +
  annotate("text", x = params_cdi[params_cdi$number.ws == 508, "b"] - 0.05, y = params_cdi[params_cdi$number.ws == 508, "a"] + 0.15, colour = "black", label = "(93.3%)") +
  xlab("Difficulty") +
  ylab("Discrimination") +
  labs(colour = "Frequency (%)") +
  scale_color_gradient(low="green", high="red") + 
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

Most popular categories:

```{r}
cat_categories <- data.frame(matrix(NA, nrow = length(sim_results), ncol = length(unique(cdi$category))))
colnames(cat_categories) <- unique(cdi$category)

n <- 1
for (sim in sim_results){
  person_items <- sim$items_answered
  person_cats <- table(cdi[person_items, "category"]) / length(person_items)
  for (col in colnames(cat_categories)){
    if (col %in% names(person_cats)){
      cat_categories[[col]][n] <- person_cats[[col]]
    } else {
      cat_categories[[col]][n] <- 0
    }
  }
  n <- n + 1
}

cat_categories <- data.frame(category = unique(cdi$category), freq = round( apply(cat_categories, 2, function(x) mean(x)), 3) * 100)
cdi_categories <- as.data.frame(round(table(cdi$category) / nrow(cdi), 3) * 100)
colnames(cdi_categories) <- c("category", "freq")
categories_used <- merge(cat_categories, cdi_categories, by="category", suffixes = c("_CAT", "_full"))
#categories_used$diff <- categories_used$freq_CAT - categories_used$freq_full
categories_used$mean_dscr <- round(aggregate(a ~ category, data = params_cdi, mean)$a, 2)
categories_used$mean_dfclt <- round(aggregate(b ~ category, data = params_cdi, mean)$b, 2)
categories_used
```

```{r eval=F, include=F}
#for thesis
write.table(categories_used, file="categories.txt", quote = F, row.names = F)
```

