---
title: "Model fit"
author: "Piotr Król"
date: "18 08 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Content:

* Full model
  + Items fit
  + Local dependence
* Restricted model
  + Items fit
  + Local dependence

```{r libraries, warning=FALSE, message=FALSE}
library(mirt)
library(DescTools) #For normal Cramer's V calculation
```

Load already prepared data:

```{r data loading, eval=F}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
cdi <- read.csv("Data/cdi.csv", encoding = "UTF-8")
```

Speed up some computations:

```{r eval=F}
mirtCluster()
```

```{r ready report data, include=F}
load("Data/model_fit.RData")
```

# Full model

```{r eval=F, include=F}
mod <- mirt(data = responses, model = 1, SE = TRUE, quadpts = 601, technical = list(NCYCLES = 10000))
save(mod, file="Data/mod")
```

```{r}
mod
```

## Items fit

```{r eval=F}
itemfit <- itemfit(mod, method = "MAP") #Use default S_X2 statistic and MAP ability estimation
```

```{r}
head(itemfit)
```

Misfits for p = 0.01:

```{r}
length(which(itemfit$p.S_X2 < 0.01))
```
```{r}
itemfit[itemfit$p.S_X2 < 0.01, ]
```

```{r}
cdi[row.names(itemfit[itemfit$p.S_X2 < 0.01, ]), ]
```

Misfits for p = 0.001:

```{r}
length(which(itemfit$p.S_X2 < 0.001))
```
```{r}
itemfit[itemfit$p.S_X2 < 0.001, ]
```

```{r}
cdi[row.names(itemfit[itemfit$p.S_X2 < 0.001, ]), ]
```

```{r}
hist(itemfit$RMSEA.S_X2)
```

## Local dependence

### Residuals 

"By default, residuals() computes the local dependence (LD)
pairwise statistic between each pair of items, which is very similar to a signed 2 value (Chen
and Thissen 1997). Also, a standardized version of the LD statistic (Cramer's V) is printed
above the diagonal to aid in interpretation when items contain more than two response options
and hence more degrees of freedom" ~ mirt documentation

```{r eval=F}
residuals <- residuals(mod)
```

```{r}
residuals_up <- residuals
residuals_down <- residuals
residuals_up[lower.tri(residuals_up)] <- NA
residuals_down[upper.tri(residuals_down)] <- NA
```

```{r}
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramers V")
```
```{r}
summary(as.vector(residuals_up))
```
Check positions with the highest Cramer's V:

```{r}
indexes <- arrayInd(which(residuals_up > 0.3), dim(residuals))
data.frame(pos1 = cdi[indexes[, 1], "position"], pos2 = cdi[indexes[, 2], "position"], CramersV = residuals_up[indexes])
```
Check positions with the lowest Cramer's V:

```{r}
indexes <- arrayInd(which(residuals_up < -0.14), dim(residuals))
data.frame(pos1 = cdi[indexes[, 1], "position"], pos2 = cdi[indexes[, 2], "position"], CramersV = residuals_up[indexes])
```

```{r}
hist(residuals_down, main="Histogram of Chi-squares", xlab= "Chi-square")
```
```{r}
summary(as.vector(residuals_down))
```
Check positions with the highest Chi-squares:

```{r}
indexes <- arrayInd(which(residuals_down > 200), dim(residuals))
data.frame(pos1 = cdi[indexes[, 1], "position"], pos2 = cdi[indexes[, 2], "position"], Chi.sq = residuals_down[indexes])
```

How many items have Chi.sq > 100 with at least one item:

```{r}
indexes <- arrayInd(which(residuals_down > 100), dim(residuals))
rows <- indexes[, 1]
columns <- indexes[, 1]
unique <- unique(c(unique(rows), unique(columns)))
length(unique)
```

Check positions with the lowest Chi-squares:

```{r}
indexes <- arrayInd(which(residuals_down < 0.009), dim(residuals))
data.frame(pos1 = cdi[indexes[, 1], "position"], pos2 = cdi[indexes[, 2], "position"], Chi.sq = residuals_down[indexes])
```

### Normal Cramer's V

Check normal Cramer's V values:

```{r eval=F}
cramersV <- PairApply(responses, FUN = CramerV)
cramersV[lower.tri(cramersV)] <- NA
diag(cramersV) <- NA
```

```{r}
hist(cramersV, main="Histogram of Cramer's V values", xlab= "Cramer's V")
```
How many items correlate higher than 0.8 with at least one item:

```{r}
indexes <- arrayInd(which(cramersV > 0.8), dim(cramersV))
rows <- indexes[, 1]
columns <- indexes[, 1]
unique <- unique(c(unique(rows), unique(columns)))
length(unique)
```

```{r}
indexes <- arrayInd(which(cramersV > 0.8), dim(cramersV))
data.frame(pos1 = cdi[indexes[, 1], "position"], pos2 = cdi[indexes[, 2], "position"], cramersV = cramersV[indexes])
```

# Restricted model

```{r}
items_to_remove <- as.numeric(row.names(itemfit[itemfit$p.S_X2 < 0.001, ]))
items_to_remove
```

```{r eval=F, include=F}
save(items_to_remove, file="Data/items_to_remove")
```

Model created on restricted set of items:

```{r eval=F, include=F}
mod_r <- mirt(data = responses[, -c(392, 417, 652)], model = 1, SE = TRUE, quadpts = 601, technical = list(NCYCLES = 10000))
save(mod_r, file="Data/mod_r")
```
```{r}
mod_r
```
## Items fit

```{r eval=F}
itemfit_r <- itemfit(mod_r, method = "MAP")
```

Check if there are any misfits for p < 0.001:

```{r}
length(which(itemfit_r$p.S_X2 < 0.001))
```
```{r}
cdi[row.names(itemfit_r[itemfit_r$p.S_X2 < 0.001, ]), ]
```
```{r}
itemfit_r[itemfit_r$p.S_X2 < 0.001, ]
```
## Local dependence

```{r eval=F}
residuals_r <- residuals(mod_r)
```

```{r}
residuals_up <- residuals_r
residuals_down <- residuals_r
residuals_up[lower.tri(residuals_up)] <- NA
residuals_down[upper.tri(residuals_down)] <- NA
```

```{r}
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramers V")
```

```{r}
summary(as.vector(residuals_up))
```