---
title: "IRMiK: Słowa i Gesty - vocabulary comprehension"
author: "Piotr Król"
date: "07 10 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Load libraries and functions:

```{r libraries, warning=FALSE, message=FALSE}
library(mirt)
library(mirtCAT)
library(parallel)
library(plyr)
library(ggpubr)
source(paste0(getwd(),"/Functions/prepare_data.R"))
source(paste0(getwd(),"/Functions/get_sim_results.R"))
```

Load data downloaded from https://drive.google.com/file/d/1i4YupJCfE6qt1iZ9A6iqZaxXv71gZ_JV/view 

```{r}
data <- read.csv("Data/irmik_lex.csv", encoding = "UTF-8")
names(data)[names(data) == 'Płeć'] <- 'Plec' #Polish symbols causes some problems later when preparing data
```

# Data preparation & data characteristics

Prepare data with ready function:

```{r eval=F}
prepare_data(data, "G", F)
rm(d)
```

```{r include=F}
load("Data/SiG_comp.RData")
```

Prepare responses matrix

```{r}
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
```

Number of respondents:

```{r}
nrow(responses)
```

Number of items:

```{r}
ncol(responses)
```

Number of floor respondents:

```{r}
length(which(rowSums(responses) == 0))
```

Number of ceiling respondents:

```{r}
length(which(rowSums(responses) == 380))
```
# All children (8 - 18 months)

## Model creation & quadratures

301 quadratures

```{r eval=F}
mod301q <- mirt(data = responses, model = 1, SE = TRUE, quadpts = 301, technical = list(NCYCLES = 30000))
```

```{r}
mod301q
```

451 quadratures

```{r eval=F}
mod451q <- mirt(data = responses, model = 1, SE = TRUE, quadpts = 451, technical = list(NCYCLES = 30000))
```

```{r}
mod451q
```

I select model created with use of **451 quadrature points** for further analysis:

```{r}
mod <- mod451q
```

## Model fit

### Items fit

```{r eval=F}
itemfit <- itemfit(mod, method = "MAP")
```

```{r}
itemfit[itemfit$p.S_X2 < 0.001, ]
```

```{r}
cdi[which(cdi$number.wg == 104), ]
```

### Local dependence

```{r eval=F}
residuals <- residuals(mod)
```

```{r}
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

```{r}
summary(as.vector(residuals_up))
```

The most highly dependent positions:

```{r}
indexes <- arrayInd(which(residuals_up > 0.3), dim(residuals))
data.frame(pos1 = cdi[indexes[, 1], "position"], pos2 = cdi[indexes[, 2], "position"], CramersV = residuals_up[indexes])
```

## Model creation once more

Remove one misfit and create model once more:

```{r eval=F}
responses_r <- responses[, -which((colnames(responses) %in% c("item104")))]
cdi_r <- cdi[cdi$number.wg != 104, ]
mod_r <- mirt(data = responses_r, model = 1, SE = TRUE, quadpts = 451, technical = list(NCYCLES = 30000))
```

### Items fit

```{r eval=F}
itemfit_r <- itemfit(mod_r, method = "MAP")
```

```{r}
itemfit_r[itemfit_r$p.S_X2 < 0.001, ]
```
No misfits now. Hurra!

## Items parameteres

```{r eval=F}
params_IRT <- as.data.frame(coef(mod_r, simplify = T, IRTpars = T)$items)[1:2]
params_cdi <- cbind(params_IRT, cdi_r)
```

```{r fig.width=12, fig.height=8}
ggplot(params_cdi, aes(b, a, colour=category, label=position)) +
  geom_text(check_overlap = T) +
  geom_point(alpha = 0.3) +
  xlab("Difficulty") +
  ylab("Discrimination") +
  labs(colour = "Category") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

## Ability vs. SE

```{r eval=F}
fscores <- as.data.frame(fscores(mod_r, method = "MAP", full.scores.SE = T))
```

```{r fig.width=12, fig.height=8}
ggplot(fscores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Standard error (SE)") +
  labs(title = "CDI: WG comprehension (all children)") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

## Fixed-length simulations

```{r eval=F}
params <- as.data.frame(coef(mod_r, simplify = T)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
lengths <- c(5, 10, 15, 25, 35, 50, 75, 100)
results_fixed_length <- data.frame(cor = NA, meanSE = NA, rel = NA, unused = NA)
cl <- makeCluster(detectCores())

for (n in lengths){
  
  #Make simulation
  sim_results <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_items = n, max_items = n))
  
  #Obtain thetas
  thetas <- laply(sim_results, function(x) x$thetas)
  
  #Get correlation of thetas with full scores
  cor <- round(cor(thetas, fscores$F1), 3)
  
  #Get mean SE
  meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)
  
  #Get reliability
  rel <- round(1 - meanSE**2, 3)
  
  #Get number of unused items
  items_answered <- laply(sim_results, function(x) x$items_answered)
  unused <- nrow(params) - length(unique(as.vector(items_answered)))
  
  #Update data frame with results
  results_fixed_length <- rbind(results_fixed_length, c(cor, meanSE, rel, unused))
  
}

results_fixed_length <- na.omit(results_fixed_length)
row.names(results_fixed_length) <- lengths
```

```{r}
results_fixed_length
```

## Simulation with SE = 0.1 stop criterion

```{r eval=F}
results_SE1 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1))
```

```{r}
get_sim_results(results_SE1, fscores, params)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE1, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 50, 75, 100, 378, 379), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 50", "51 - 75", "76 - 100", "101 - 378", "379"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  labs(title = "CDI: WG comprehension (all children). SE = 0.1 stop criterion") +
  theme_pubclean() +
  theme(text = element_text(size=16))
```

## Simulation with SE = 0.15 stop criterion

```{r eval=F}
results_SE15 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.15))
```

```{r}
get_sim_results(results_SE15, fscores, params)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE15, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 50, 75, 100, 378, 379), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 50", "51 - 75", "76 - 100", "101 - 378", "379"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  labs(title = "CDI: WG comprehension (all children). SE = 0.15 stop criterion") +
  theme_pubclean() +
  theme(text = element_text(size=16))
```

## SE vs. age

```{r fig.width=12, fig.height=8}
ggplot(data.frame(age = as.factor(responses_demo$months), SE = fscores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab("Standard error (SE)") +
  labs(title = "CDI: WG comprehension") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
  )
```

# 12 - 18 months children

```{r}
responses_demo2 <- responses_demo[responses_demo$months >= 12, ]
responses2 <- as.matrix(responses_demo2[,5:ncol(responses_demo2)])
```

## SE vs. age

```{r}
fscores2 <- fscores[which(responses_demo$months >= 12), ]
```

```{r fig.width=12, fig.height=8}
ggplot(fscores2, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Standard error (SE)") +
  labs(title = "CDI: WG comprehension (12 - 18 months children)") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

## Simulation with SE = 0.1 stop criterion

```{r eval=F}
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
results_SE1_2 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses2, cl = cl, design = list(min_SEM = 0.1))
```

```{r}
get_sim_results(results_SE1_2, fscores2, params)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE1_2, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 50, 75, 100, 378, 379), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 50", "51 - 75", "76 - 100", "101 - 378", "379"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses2) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  labs(title = "CDI: WG comprehension (12 - 18 months children). SE = 0.1 stop criterion") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

## Simulation with SE = 0.15 stop criterion

```{r eval=F}
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
results_SE15_2 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses2, cl = cl, design = list(min_SEM = 0.15))
```

```{r}
get_sim_results(results_SE15_2, fscores2, params)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE15_2, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 50, 75, 100, 378, 379), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 50", "51 - 75", "76 - 100", "101 - 378", "379"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses2) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  labs(title = "CDI: WG comprehension (12 - 18 months children). SE = 0.15 stop criterion") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```