---
title: "IRMiK: Słowa i Gesty - gestures"
author: "Piotr Król"
date: "24 01 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Load libraries and functions:

```{r libraries, warning=FALSE, message=FALSE}
library(mirt)
library(mirtCAT)
library(parallel)
library(plyr)
library(ggpubr)
source(paste0(getwd(),"/Functions/get_sim_results.R"))
source(paste0(getwd(),"/Functions/get_first_item_responses.R"))
source(paste0(getwd(),"/Functions/misfits_removal.R"))
```

# Data preparation

```{r include=F}
load("Data/SiG_gestures.RData")
```

Load data sent by Grzegorz and prepared responses from SiG comprehension:

```{r eval=F}
data <- read.csv("Data/gesty_pozycje.csv", encoding = "UTF-8")
load("Data/responses_demo_comp")
responses_demo_comp <- responses_demo
rm(responses_demo)
```

```{r}
data <- data[order(data$KOD), ]
```

Check if codes from the two sets are the same:

```{r}
setequal(responses_demo_comp$participant, data$KOD)
```

Prepare df with gestures answers only :

```{r eval=F}
columns_to_rm <- c("KOD")
for (letter in c("A", "B", "C", "D", "E")){
  columns_to_rm <- c(columns_to_rm,
                       sprintf("Gesty_%s_SUMA_NIE", letter),
                       sprintf("Gesty_%s_SUMA_TAK", letter),
                       sprintf("Gesty_%s_Uwagi", letter)
                     )
}
data_pure <- data[, !(names(data) %in% columns_to_rm)]
```

Check if number of gestures is correct:

```{r}
ncol(data_pure) / 2
```
Prepare gestures responses data frame:

```{r eval=F}
#Prepare empty responses df with proper names
responses_demo <- data.frame(matrix(NA, nrow = nrow(data_pure), ncol = ncol(data_pure) / 2))
gestures_names <- c()
for (name in names(data_pure)){
  gestures_names <- c(gestures_names, substr(name, 1, nchar(name) - 4))
}
gestures_names <- unique(gestures_names)
names(responses_demo) <- gestures_names

#Fill in responses df with 0s and 1s
i <- 1
j <- 2

for (c in 1:(ncol(data_pure) / 2)){
  cat(paste0(c, " gestures of ", ncol(data_pure) / 2, " prepared. \r"))
  for (r in 1:nrow(data_pure[, i:j])){
    answer <- data_pure[r, i:j]
    if(!is.na(answer[1]) & !is.na(answer[2])){
      if (answer[1] == 0 & answer[2] == 1) responses_demo[r, c] = 1
      if ((answer[1] == 1 & answer[2] == 0) | (answer[1] == 0 & answer[2] == 0)) responses_demo[r, c] = 0
    }
  }
  i <- i + 2
  j <- j + 2
}

rm(answer)

#Add demo columns from WG comp responses df
responses_demo$participant <- data$KOD
responses_demo$days <- responses_demo_comp$days
responses_demo$months <- responses_demo_comp$months
responses_demo$gender <- responses_demo_comp$gender

#Change column ordering
col_order <- c("participant", "days", "months", "gender", names(responses_demo)[1:63])
responses_demo <- responses_demo[, col_order]

#Remove 20 respondents who gave no answers to gestures
responses_demo0 <- responses_demo #Keep variable
responses_demo <- na.omit(responses_demo)
```

# Model creation

Prepare responses matrix for model creation:

```{r}
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
```

## Quadrature points

### 61 quadrature points

```{r eval=F}
mod61q <- mirt(data = responses, model = 1, SE = TRUE, quadpts = 61)
```

```{r}
mod61q
```

### 101 quadrature points

```{r eval=F}
mod101q <- mirt(data = responses, model = 1, SE = TRUE, quadpts = 101)
```

```{r}
mod101q
```
The same log-likelihood. It means that 61 quadrature points is optimal value

## Items fit

Run script for removing misfits:

```{r eval = F}
output <- misfits_removal(responses=responses, quadtps=61, NCYCLES=1000, p=0.001)
mod <- output[[1]]
items_removed <- output[[2]]
```

Check what items were removed:

```{r}
items_removed
```
**Interesting: all the removed items come from the 'E' category**

Remove these items from responses matrix:

```{r}
responses0 <- responses #Store old matrix just in case
responses <- responses[, -which((colnames(responses) %in% unlist(items_removed)))]
print(paste0(ncol(responses), " items left after removal."))
```

## Local dependence

```{r eval=F}
residuals <- residuals(mod)
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA
```

```{r}
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

```{r}
summary(as.vector(residuals_up))
```

No need to remove items by means of local dependence

# Model exploration

## IRT items parameters

```{r fig.width=12, fig.height=8}
params_IRT <- as.data.frame(coef(mod, simplify = T, IRTpars = T)$items)[1:2]
params_IRT$category <- substr(row.names(params_IRT), 7, 7)

ggplot(params_IRT, aes(b, a, colour=category)) +
  geom_point(alpha = 1) +
  xlab("Difficulty") +
  ylab("Discrimination") +
  labs(colour = "Gestures category") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

## Ability vs. SE

### All children

```{r eval=F}
fscores <- as.data.frame(fscores(mod, method = "MAP", full.scores.SE = T))
```

```{r fig.width=12, fig.height=8}
ggplot(fscores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Standard error (SE)") +
  labs(title = "Gestures (all children)") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

```{r}
summary(fscores$SE_F1)
```
### 16-18 months children

```{r fig.width=12, fig.height=8}
fscores_r <- fscores[which(responses_demo$months >= 16),]
ggplot(fscores_r, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Standard error (SE)") +
  labs(title = "Gestures (16 - 18 months children)") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

```{r}
summary(fscores_r$SE_F1)
```

## SE vs. age

```{r fig.width=12, fig.height=8 }
ggplot(data.frame(age = as.factor(responses_demo_comp[responses_demo_comp$participant %in% responses_demo$participant, ]$months), SE = fscores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab("Standard error of ability estimations") +
  labs(title = "Gestures") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
  )
```

# Simulations (only 16 - 18 months children)

```{r eval=F}
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
```

Get only 16-18 months children:

```{r}
responses_r <- responses[which(responses_demo$months >= 16), ]
nrow(responses_r)
```

## Whole pool for everyone

```{r eval=F}
results_all <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r, cl = cl, design = list(min_items = ncol(responses_r)))
```

```{r}
get_sim_results(results_all, fscores_r, params)
```

```{r fig.width=12, fig.height=8}
SE_history <- laply(results_all, function(x) x$thetas_SE_history)
meanSE_item <- as.data.frame(colMeans(SE_history))
meanSE_item$item <- row.names(meanSE_item)
colnames(meanSE_item) <- c("mean_SE", "item")
meanSE_item$item <- as.numeric(meanSE_item$item)

ggline(meanSE_item, x = "item", y = "mean_SE", numeric.x.axis = TRUE) +
  xlab ("Test length") +
  ylab ("Mean SE") +
  labs(title = "Gestures (16 - 18 months children)") +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  ) + 
  scale_y_continuous(breaks = c(0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)) + 
  scale_x_continuous(breaks = seq(0, nrow(params), by=5), limits = c(1, nrow(params)))
```

## SE = 0.2 stop criterion

```{r eval=F}
results_SE2 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r, cl = cl, design = list(min_SEM = 0.2))
```

```{r}
get_sim_results(results_SE2, fscores_r, params)
```

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE2, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 45, 54, 55), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 45", "46-54", "55 (all)"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses_r) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  labs(title = "Gestures (16 - 18 months children). SE = 0.2 stop criterion") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

## Different starting thetas

Prepare full scores with demographic variables:

```{r}
fscores_demo <- cbind(responses_demo[, c("days", "months", "gender")], fscores)
fscores_demo$weeks <- ceiling(fscores_demo$days / 7)
fscores_demo <- fscores_demo[which(fscores_demo$months >= 16), ]
nrow(fscores_demo)
```

Get full scores aggregated:

```{r}
fscores_aggr <- aggregate(F1 ~ months + gender, data = fscores_demo, mean)
head(fscores_aggr)
```

Get start theta for each respondent by months and gender:

```{r}
age_unit <- colnames(fscores_aggr)[1]
start_thetas <- as.matrix(
    mapply(
      function(age, gender){
        return(fscores_aggr[fscores_aggr[[age_unit]] == age & fscores_aggr$gender == gender, "F1"])
      },
      fscores_demo[[age_unit]],
      fscores_demo$gender
    )
  )
length(start_thetas)
```

### No start theta given

```{r}
get_sim_results(results_SE2, fscores_r, params)
```

```{r}
get_first_item_responses(results_SE2, responses_r)
```
### Start theta by age and gender

```{r eval=F}
results_SE2_theta <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r, cl = cl, design = list(min_SEM = 0.2, thetas.start = start_thetas))
```

```{r}
get_sim_results(results_SE2_theta, fscores_r, params)
```

```{r}
get_first_item_responses(results_SE2_theta, responses_r)
```

### Start thetas decreased by 2

```{r eval=F}
results_SE2_theta2 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r, cl = cl, design = list(min_SEM = 0.2, thetas.start = start_thetas - 2))
```

```{r}
get_sim_results(results_SE2_theta2, fscores_r, params)
```

```{r}
get_first_item_responses(results_SE2_theta2, responses_r)
```

# Prepare input files

```{r}
items <- params
items$position <- row.names(params)
items$question <- "Czy Twoje dziecko wykonuje"
items$group <- "gesty"
write.csv(items, file = "items-WG-gestures.csv", fileEncoding = "UTF-8", row.names = F)
```

```{r}
start_thetas_df <- fscores_aggr
start_thetas_df$F1 <- start_thetas_df$F1 - 2
colnames(start_thetas_df) <- c("age", "gender", "theta")
start_thetas_df$group <- "gesty"
write.csv(start_thetas_df, file = "startThetas-WG-gestures.csv", fileEncoding = "UTF-8", row.names = F)
```

