---
title: "IRMiK: Słowa i Gesty - gestures"
author: "Piotr Król"
date: "21 10 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Load libraries and functions:

```{r libraries, warning=FALSE, message=FALSE}
library(mirt)
library(mirtCAT)
library(parallel)
library(plyr)
library(ggpubr)
source(paste0(getwd(),"/Functions/get_sim_results.R"))
```

Load data sent by Grzegorz and prepared responses from SiG comprehension:

```{r eval=F}
data <- read.csv("Data/gesty_pozycje.csv", encoding = "UTF-8")
load("Data/responses_demo_comp")
responses_demo_comp <- responses_demo
rm(responses_demo)
```

```{r include=F}
load("Data/SiG_gestures.RData")
```

# Data preparation & data characteristics

```{r eval=F}
data <- data[order(data$KOD), ]
```

Check if codes from the two sets are the same:

```{r}
setequal(responses_demo_comp$participant, data$KOD)
```
Prepare df with gestures answers only :

```{r eval=F}
columns_to_rm <- c("KOD")
for (letter in c("A", "B", "C", "D", "E")){
  columns_to_rm <- c(columns_to_rm,
                       sprintf("Gesty_%s_SUMA_NIE", letter),
                       sprintf("Gesty_%s_SUMA_TAK", letter),
                       sprintf("Gesty_%s_Uwagi", letter)
                     )
}
data_pure <- data[, !(names(data) %in% columns_to_rm)]
```

```{r}
ncol(data_pure) / 2
```
It's equal to the number of gestures in SiG.

Prepare proper responses data frame:

```{r eval=F}
#Prepare empty responses df with proper names
responses_demo <- data.frame(matrix(NA, nrow = nrow(data_pure), ncol = ncol(data_pure) / 2))
gestures_names <- c()
for (name in names(data_pure)){
  gestures_names <- c(gestures_names, substr(name, 1, nchar(name) - 4))
}
gestures_names <- unique(gestures_names)
names(responses_demo) <- gestures_names

#Fill in responses df with 0s and 1s
i <- 1
j <- 2

for (c in 1:(ncol(data_pure) / 2)){
  cat(paste0(c, " gestures of ", ncol(data_pure) / 2, " prepared. \r"))
  for (r in 1:nrow(data_pure[, i:j])){
    answer <- data_pure[r, i:j]
    if(!is.na(answer[1]) & !is.na(answer[2])){
      if (answer[1] == 0 & answer[2] == 1) responses[r, c] = 1
      if ((answer[1] == 1 & answer[2] == 0) | (answer[1] == 0 & answer[2] == 0)) responses_demo[r, c] = 0
    }
  }
  i <- i + 2
  j <- j + 2
}

rm(answer)
responses_demo$participant <- data$KOD
```

```{r}
nrow(responses_demo)
```

```{r eval=F}
responses_demo_noNA <- na.omit(responses_demo)
```

```{r}
nrow(responses_demo_noNA)
```

** 20 respondents were removed because of no answers for gestures **

Prepare responses matrix for further IRT and CAT analyses

```{r eval=F}
responses <- responses_demo_noNA
responses$participant <- NULL
responses <- as.matrix(responses)
```

Number of floor respondents:

```{r}
length(which(rowSums(responses) == 0))
```
Number of ceiling respondents:

```{r}
length(which(rowSums(responses) == 63))
```

# Model creation

## 61 quadratures

```{r eval=F}
mod61q <- mirt(data = responses, model = 1, SE = TRUE, quadpts = 61)
```

```{r}
mod61q
```

## 101 quadratures

```{r eval=F}
mod101q <- mirt(data = responses, model = 1, SE = TRUE, quadpts = 101)
```

```{r}
mod101q
```
Yay, no need to increase quadratures.

I select model created with use of **default 61 quadrature points** for further analysis:

```{r}
mod <- mod61q
```

# Model fit

## Items fit

```{r}
itemfit <- itemfit(mod, method = "MAP")
itemfit[itemfit$p.S_X2 < 0.001, ]
```

```{r}
itemfit <- itemfit(mod61q, method = "MAP")
itemfit[itemfit$p.S_X2 < 0.05, ]
```

Two misfits. Let's remove them and create model once more:

```{r}
responses_r <- responses[, !colnames(responses) %in% itemfit[itemfit$p.S_X2 < 0.001, "item"]]
ncol(responses_r)
```

```{r eval=F}
mod_r <- mirt(data = responses_r, model = 1, SE = TRUE, quadpts = 61)
```

```{r }
itemfit <- itemfit(mod_r, method = "MAP")
itemfit[itemfit$p.S_X2 < 0.001, ]
```

Theoretically I should remove it and create model once more, but when I started this process I could not end (ca 5 models created and still misfits).

```{r}
mod <- mod_r
```

## Local dependence

```{r eval=F}
residuals <- residuals(mod)
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA
```

```{r}
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

```{r}
summary(as.vector(residuals_up))
```
No need to remove anything by means of local dependence.

# Items parameters

```{r fig.width=12, fig.height=8}
params_IRT <- as.data.frame(coef(mod, simplify = T, IRTpars = T)$items)[1:2]
params_IRT$category <- substr(row.names(params_IRT), 7, 7)

ggplot(params_IRT, aes(b, a, colour=category)) +
  geom_point(alpha = 1) +
  xlab("Difficulty") +
  ylab("Discrimination") +
  labs(colour = "Category") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

# Ability vs. SE

```{r eval=F}
fscores <- as.data.frame(fscores(mod, method = "MAP", full.scores.SE = T))
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2]
```

```{r fig.width=12, fig.height=8}
ggplot(fscores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Standard error (SE)") +
  labs(title = "Gestures (all children)") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

# SE vs. age

```{r fig.width=12, fig.height=8 }
ggplot(data.frame(age = as.factor(responses_demo_comp[responses_demo_comp$participant %in% responses_demo_noNA$participant, ]$months), SE = fscores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab("Standard error of ability estimations") +
  labs(title = "Gestures") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
  )
```

# Simulation with SE = 0.15 stop criterion

```{r eval=F}
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
results_SE15 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r, cl = cl, design = list(min_SEM = 0.15))
```

```{r}
get_sim_results(results_SE15, fscores, params)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE15, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 45, 60, 61), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 45", "46-60", "61"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses_r) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  labs(title = "Gestures (all children). SE = 0.15 stop criterion") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

# Simulation with SE = 0.2 stop criterion

```{r eval=F}
results_SE2 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r, cl = cl, design = list(min_SEM = 0.2))
```

```{r}
get_sim_results(results_SE2, fscores, params)
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE2, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 45, 60, 61), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 45", "46-60", "61"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses_r) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  labs(title = "Gestures (all children). SE = 0.2 stop criterion") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

# Fixed-length simulations

```{r eval=F}
lengths <- c(5, 10, 15, 25, 35, 45)
results_fixed_length <- data.frame(cor = NA, meanSE = NA, rel = NA, unused = NA)

for (n in lengths){
  
  #Make simulation
  sim_results <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r, cl = cl, design = list(min_items = n, max_items = n))
  
  #Obtain thetas
  thetas <- laply(sim_results, function(x) x$thetas)
  
  #Get correlation of thetas with full scores
  cor <- round(cor(thetas, fscores$F1), 3)
  
  #Get mean SE
  meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)
  
  #Get reliability
  rel <- round(1 - meanSE**2, 3)
  
  #Get number of unused items
  items_answered <- laply(sim_results, function(x) x$items_answered)
  unused <- nrow(params) - length(unique(as.vector(items_answered)))
  
  #Update data frame with results
  results_fixed_length <- rbind(results_fixed_length, c(cor, meanSE, rel, unused))
  
}

results_fixed_length <- na.omit(results_fixed_length)
row.names(results_fixed_length) <- lengths
```

```{r}
results_fixed_length
```

# 12-18 months children. SE = 0.15 stop criterion

```{r}
responses_r2 <- responses_demo[which(responses_demo_comp$months >= 12), ]
responses_r2 <- as.matrix(na.omit(responses_r2))
nrow(responses_r2)
```

```{r eval=F}
results_SE15_2 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r2, cl = cl, design = list(min_SEM = 0.15))
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE15_2, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 45, 60, 61), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 45", "46-60", "61"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses_r2) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  labs(title = "Gestures (12 - 18 months). SE = 0.15 stop criterion") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

# 12-18 months children. SE = 0.2 stop criterion

```{r eval=F}
results_SE2_2 <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses_r2, cl = cl, design = list(min_SEM = 0.2))
```

Plot distribution of tests lengths:

```{r fig.width=12, fig.height=8}
#Prepare cuts
tests_lengths <- laply(results_SE2_2, function(x) length(x$items_answered))
cuts <- cut(tests_lengths, breaks = c(0, 5, 10, 15, 25, 35, 45, 60, 61), labels = c("1 - 5", "6 - 10", "11 - 15", "16 - 25", "26 - 35", "36 - 45", "46-60", "61"))

#Plot
ggplot(data.frame(round(table(cuts) / nrow(responses_r2) * 100, 1)), aes(x = cuts, y = Freq)) +
  xlab("Number of items administered") +
  ylab("Percent of respondents (%)") +
  geom_bar(stat = "identity") +
  labs(title = "Gestures (12 - 18 months). SE = 0.2 stop criterion") +
  geom_text(aes(label = Freq), vjust = -0.3, size=6) + 
  theme_pubclean() +
  theme(text = element_text(size=16))
```

