---
title: "Starting item"
author: "Piotr Król"
date: "15 07 2021"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

```{r libraries, warning=F, message=F}
library(mirt)
library(ggpubr)
library(parallel)
library(plyr)
```

```{r data loading, eval=F}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
load("Data/mod_r")
load("Data/items_to_remove")
responses <- responses[, -items_to_remove]
```

```{r data loading 2, include=F}
load("Data/starting_item.RData")
```

```{r model}
mod <- mod_r
rm(mod_r)
mod
```

Prepare full scores, params and mirt object for simulations:

```{r preparation, eval=F}
fscores <- as.data.frame(fscores(mod, method = "MAP", full.scores.SE = T)) #Obtain full scores
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2] #Prepare params
mo <- generate.mirt_object(params, '2PL') #Prepare mirt object
```

Prepare full scores with demographic variables:

```{r eval=F}
fscores_demo <- cbind(responses_demo[, c("days", "months", "gender")], fscores)
fscores_demo$weeks <- ceiling(fscores_demo$days / 7)
fscores_demo[fscores_demo$gender == "chłopiec", "gender"] <- "Boy"
fscores_demo[fscores_demo$gender == "dziewczynka", "gender"] <- "Girl"
```


```{r}
fscores_aggr <- aggregate(F1 ~ gender, data = fscores_demo, mean)
fscores_aggr
```

```{r}
fscores_aggr <- aggregate(F1 ~ months + gender, data = fscores_demo, mean)
ggplot(fscores_aggr, aes(months, F1, colour=gender)) +
  geom_point() +
  xlab("Age (months)") +
  ylab("Ability level (Θ)") +
  labs(colour = "Sex") +
  scale_fill_discrete(labels = c('Boy','Girl'),   guide = guide_legend(reverse=TRUE)) +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```
```{r}
cor(fscores_aggr$months, fscores_aggr$F1)
```

Get start theta for each subject by **months** and gender:

```{r}
age_unit <- colnames(fscores_aggr)[1]
start_thetas <- as.matrix(
    mapply(
      function(age, gender){
        return(fscores_aggr[fscores_aggr[[age_unit]] == age & fscores_aggr$gender == gender, "F1"])
      },
      fscores_demo[[age_unit]],
      fscores_demo$gender
    )
  )
```

Make simulation without start thetas given:

```{r eval=F}
#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
results <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10))

save.image("Data/starting_item.RData")
```

```{r}
sim_results <- results

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used

paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused:", unused)
```

```{r}
first_items <- laply(sim_results, function(x) x$items_answered[1])
first_item_responses <- mapply(
      function(subject, first_item){
        return(raw_responses[subject, first_item])
      },
      c(1:nrow(responses)),
      first_items
    )
round(table(first_item_responses) / nrow(responses), 3) #2 means selecting an item
```
```{r}
ggplot(data.frame(thetas = thetas, lengths = tests_lengths), aes(thetas, lengths)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Number of items administered") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey")
)
```

Make simulation with start thetas given by age and gender:

```{r eval=F}
#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
results_thetasGiven <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10, thetas.start = start_thetas))

save.image("Data/starting_item.RData")
```

```{r}
sim_results <- results_thetasGiven

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used

paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused:", unused)
```
```{r}
first_items <- laply(sim_results, function(x) x$items_answered[1])
first_item_responses <- mapply(
      function(subject, first_item){
        return(raw_responses[subject, first_item])
      },
      c(1:nrow(responses)),
      first_items
    )
round(table(first_item_responses) / nrow(responses), 3) #2 means selecting an item
```
```{r}
ggplot(data.frame(thetas = thetas, lengths = tests_lengths), aes(thetas, lengths)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (Θ)") +
  ylab("Number of items administered") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey")
)
```

Make simulation with decreased start thetas by 0.5:

```{r eval=F}
#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
results_thetas05Given <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10, thetas.start = start_thetas - 0.5))

save.image("Data/starting_item.RData")
```

```{r}
sim_results <- results_thetas05Given

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used

paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused:", unused)
```

```{r}
first_items <- laply(sim_results, function(x) x$items_answered[1])
first_item_responses <- mapply(
      function(subject, first_item){
        return(raw_responses[subject, first_item])
      },
      c(1:nrow(responses)),
      first_items
    )
round(table(first_item_responses) / nrow(responses), 3) #2 means selecting an item
```

Make simulation with decreased start thetas by 2:

```{r eval=F}
#Prepare cluster for faster simulation
cl <- makeCluster(detectCores())

#Make simulation
results_thetas2Given <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10, thetas.start = start_thetas - 2))

save.image("Data/starting_item.RData")
```

```{r}
sim_results <- results_thetas2Given

#Obtain mean test length
tests_lengths <- laply(sim_results, function(x) length(x$items_answered))
mean_length <- round(mean(tests_lengths), 1)

#Obtain median test length
median_length <- round(median(tests_lengths), 1)

#Obtain thetas
thetas <- laply(sim_results, function(x) x$thetas)

#Get correlation of thetas with full scores
cor <- round(cor(thetas, fscores$F1), 3)

#Get mean SE
meanSE <- round(mean(laply(sim_results, function(x) x$SE_thetas)), 3)

#Get reliability
rel <- round(1 - meanSE**2, 3)

#Get number of unused items
raw_responses <- laply(sim_results, function(x) x$raw_responses)
items_used <- length(which(apply(raw_responses, 2, function(x) any(!is.na(x)))))
unused <- nrow(params) - items_used

paste("Mean length:", mean_length, " Median length:", median_length, " Correlation:", cor, " Mean SE:", meanSE, " Reliability:", rel, " Unused:", unused)
```

```{r}
first_items <- laply(sim_results, function(x) x$items_answered[1])
first_item_responses <- mapply(
      function(subject, first_item){
        return(raw_responses[subject, first_item])
      },
      c(1:nrow(responses)),
      first_items
    )
round(table(first_item_responses) / nrow(responses), 3) #2 means selecting the first item
```

```{r}
ggplot(data.frame(thetas = thetas, lengths = tests_lengths), aes(thetas, lengths)) +
  geom_point(alpha = 0.1) +
  xlab("Ability level (Θ)") +
  ylab("Number of items administered") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey")
)
```