---
title: "Starting item"
author: "Piotr Król"
date: "23 08 2021"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Content:

* Theta vs. demographic variables
* Simulations with different starting thetas

```{r libraries, warning=F, message=F}
library(mirt)
library(mirtCAT)
library(ggpubr)
library(parallel)
library(plyr)
```

```{r data loading, eval=F}
responses_demo <- read.csv("Data/responses_demo.csv", encoding = "UTF-8")
responses <- as.matrix(responses_demo[,5:ncol(responses_demo)])
load("Data/items_to_remove")
load("Data/params")
load("Data/fscores")
responses <- responses[, -items_to_remove]
source(paste0(getwd(),"/get_sim_results.R"))
```

```{r ready report data, include=F}
load("Data/starting_item.RData")
```

# Thetas vs. demographic variables

Prepare full scores with demographic variables:

```{r eval=F}
fscores_demo <- cbind(responses_demo[, c("days", "months", "gender")], fscores)
fscores_demo$weeks <- ceiling(fscores_demo$days / 7)
fscores_demo[fscores_demo$gender == "chłopiec", "gender"] <- "Boy"
fscores_demo[fscores_demo$gender == "dziewczynka", "gender"] <- "Girl"
```

```{r}
fscores_aggr <- aggregate(F1 ~ gender, data = fscores_demo, mean)
fscores_aggr
```

```{r}
fscores_aggr <- aggregate(F1 ~ months + gender, data = fscores_demo, mean)
ggplot(fscores_aggr, aes(months, F1, colour=gender)) +
  geom_point() +
  xlab("Age (months)") +
  ylab("Ability level (Θ)") +
  labs(colour = "Gender") +
  scale_fill_discrete(labels = c('Boy','Girl'),   guide = guide_legend(reverse=TRUE)) +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

```{r}
cor(fscores_aggr$months, fscores_aggr$F1)
```

# Simulations with different starting thetas

Get start theta for each subject by months and gender:

```{r}
age_unit <- colnames(fscores_aggr)[1]
start_thetas <- as.matrix(
    mapply(
      function(age, gender){
        return(fscores_aggr[fscores_aggr[[age_unit]] == age & fscores_aggr$gender == gender, "F1"])
      },
      fscores_demo[[age_unit]],
      fscores_demo$gender
    )
  )
save(start_thetas, file = "Data/start_thetas")
```

Prepare function for obtaining 1st item responses:

```{r}
get_first_item_responses <- function(sim_results, responses){
  first_items <- laply(sim_results, function(x) x$items_answered[1])
  raw_responses <- laply(sim_results, function(x) x$raw_responses)
  first_item_responses <- mapply(
        function(subject, first_item){
          return(raw_responses[subject, first_item])
        },
        c(1:nrow(responses)),
        first_items
      )
  return(round(table(first_item_responses) / nrow(responses), 3)) #2 means selecting an item
}
```

## No starting thetas given:

```{r eval=F}
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
results <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10))
```

```{r}
get_sim_results(results, fscores, params)
```

```{r}
get_first_item_responses(results, responses)
```

## Starting thetas given by age and gender:

```{r eval=F}
results_thetasGiven <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10, thetas.start = start_thetas))
```

```{r}
get_sim_results(results_thetasGiven, fscores, params)
```

```{r}
get_first_item_responses(results_thetasGiven, responses)
```

## Starting thetas given by age and gender decreased by 0.5:

```{r eval=F}
results_thetas05Given <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10, thetas.start = start_thetas - 0.5))
```

```{r}
get_sim_results(results_thetas05Given, fscores, params)
```

```{r}
get_first_item_responses(results_thetas05Given, responses)
```

## Starting thetas given by age and gender decreased by 2:

```{r eval=F}
results_thetas2Given <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10, thetas.start = start_thetas - 2))
```

```{r}
get_sim_results(results_thetas2Given, fscores, params)
```

```{r}
get_first_item_responses(results_thetas2Given, responses)
```

---

## Starting thetas given by age only (extension w.r.t. master's thesis)

```{r}
fscores_aggr <- aggregate(F1 ~ months, data = fscores_demo, mean)
ggplot(fscores_aggr, aes(months, F1)) +
  geom_point() +
  xlab("Age (months)") +
  ylab("Ability level (Θ)") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

```{r}
cor(fscores_aggr$months, fscores_aggr$F1)
```

```{r}
start_thetas <- as.matrix(
    mapply(
      function(age){
        return(fscores_aggr[fscores_aggr[[age_unit]] == age, "F1"])
      },
      fscores_demo[[age_unit]]
    )
  )
```

```{r eval=F}
results_thetasAge <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", local_pattern = responses, cl = cl, design = list(min_SEM = 0.1, max_items = 50, min_items = 10, thetas.start = start_thetas))
```

```{r}
get_sim_results(results_thetasAge, fscores, params)
```

```{r}
get_first_item_responses(results_thetasAge, responses)
```

